### 2012-02-13 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

\begin{algorithm}
\caption{Processing of Functions}
\label{ALG_processing_of_functions}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $getFunctionDeclarationNodes()$: Function defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}. \\
    $getTypeTN()$: Function defined in Algorithm \ref{ALG_extraction_of_a_type_from_typenode}. \\
    $syntaxTree$: A reference to the root node of a syntax tree.
}
\ENSURE An associative array of function names with their types and references to the their declaration nodes.

\STATE $functionArray$ = an empty associative array
\FORALL{$functionDeclarationNode \in getFunctionDeclarationNodes(syntaxTree)$}
    \STATE $functionName = functionDeclarationNode.funcName$
    \STATE $typeNode = functionDeclarationNode.getChild(TypeNode)$
    \STATE $functionArray[functionName] = (getTypeTN(typeNode), functionDeclarationNode)$
\ENDFOR
\RETURN $functionArray$
\end{algorithmic}
\end{algorithm}

@@

Later phases of the algorithm may need to process the declarations of certain functions. Therefore, we also store references to the entire function declaration nodes, accessible using function $getFunctionDeclNode()$.

@@

Another group consists of functions $set()$, $add()$, and $get()$. Function $set()$ memorizes a given value of a specified property of a specified node. It is used to assign a node with a named value, for example, $set(someNode, color, 'red')$ will assign $someNode$ with string value $'red'$ wich can be then retrieved by function $get()$, specifying the node and the property name. For example, $get(someNode, color)$ returns $'red'$. A subsequent call of $set()$ assigning a node with a value of already assigned property will overwrite it, leaving the property with the newer value. However, a property can has several values and this can be achieved using function $add()$ with the same syntax as function $set()$ but instead of overwriting the existing property, $add()$ will append the new value to the existing ones. Then, a call of $get()$ on this node and property returns a list of all its values, preserving the order of their addition.

@@

\begin{algorithm}
\caption{Function determineExpressionType()}
\label{ALG_func_determineExpressionType}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $getFunctionReturnType()$: Function returning a type of a function. \\
    $getVariableType()$: Function returning a type of a variable. \\
    $determineOperatorType()$: Function defined in \ref{ALG_func_determineOperatorType}. \\
    $getType()$: Function returning a type of already processed expression. \\
    $exprNode$: An expression node in the syntax tree. \\
    $contextVarTypes$: Types of local variables valid in the current subtree.
}
\ENSURE The expression type.

@@

\begin{algorithm}
\caption{Function determineOperatorType()}
\label{ALG_func_determineOperatorType}
\begin{algorithmic}[1]
\REQUIRE{\ \\ 
    $getType()$: Function returning a type of already processed expression. \\
    $isOperatorClassComparison()$: Function defined in \ref{ALG_func_isOperatorClassComparison}. \\
    $isOperatorClassAddition()$: Function defined in \ref{ALG_func_isOperatorClassAddition}. \\
    $isNumericType()$: Function defined in \ref{ALG_func_isNumericType}. \\
    $selectMoreGenericNumericType()$: Function that select a more generic numeric XSD type from two given types according to the type promotions rules (TODO-REF). \\
    $operatorNode$: An operator expression node in the syntax tree. \\
    $contextVarTypes$: Types of local variables valid in the current subtree.
}
\ENSURE The operator expression type.

\STATE $type = null$
\STATE $operator = operatorNode.operator$
\IF{$isOperatorClassComparison(operator)$}
    \STATE $type = xs:boolean$, exactly-one cardinality
\ELSIF{$isOperatorClassAddition(operator)$}
    \STATE $leftOperandType = getType(exprNode.leftSide)$
    \STATE $rightOperandType = getType(exprNode.rightSide)$
    \IF{$isNumericType(leftOperandType)) AND isNumericType(rightOperandType)$}
        \STATE $type = selectMoreGeneralNumericType(leftOperandType, rightOperandType)$
    \ELSIF{$isNumericType(leftOperandType)$}
        \STATE $type = leftOperandType$
    \ELSIF{$isNumericType(rightOperandType)$}
        \STATE $type = rightOperandType$
    \ENDIF
\ENDIF

@@

The algorithm iterates through the variable declaration nodes from the prolog and if the variable is explicitly assigned with its type, the type is noted. Otherwise, an attempt of the type deduction of the binding expression is made. The deduction of the type from the expression is presented in Algorithm \ref{ALG_func_determineExpressionType}. On input, it takes an expression node and types of local variables that are valid in the expression (also called context of variables or variable context). The presented algorithm is just shortened illustration since the complete version is too long to be presented in this text. Nevertheless, the principle is straightforward. Depending on the class of the expression node (show in Figure \ref{FIG_syntax_tree_node_types_part_4}), its type can be either determined directly or it depends on its subexpressions. A part of the function's semantic is separated in \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.

Function $determineExpressionType()$ is called within function $analysisOfExpressionTypes$ presented in Algorithm \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions. Two important questions about this function emerge. How (in which order) are the tree nodes recursively processed and how does the function handle definitions (bindings) of new variables which may also appear in the expressions?

The order of the recursion is to firstly process children of a node and then the node itself. The order of children processing is specified by function $o$ from the definition of the syntax tree \ref{DEF_syntax_tree}. The reason is that an expression node may need to know types of its subexpressions to determine its own type and we want to determine types of all expressions. Therefore, using the described order, subexpressions of an expression are processed first, and then the expression itself without any need for further recursion, because the types of the subexpressions are already determined.

@@

\begin{algorithm}
\caption{Function analysisOfExpressionTypes}
\label{ALG_func_analysisOfExpressionTypes}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $memorizeType()$: Function memorizing a determined type of a specified node. \\
    $set()$: Function memorizing a specified property of a specified node. \\
    $add()$: Function adding a value to a list of specified properties of a specified node. \\
    $get()$: Function returning a specified property (or list) set by functions $set()$ and $add()$. \\
    $mergeContextVarTypes()$: Function defined in Algorithm \ref{ALG_function_mergeContextVarTypes}. \\
    $startingNode$: A node determining a subtree to perform the analysis on. \\
    $contextVarTypes$: Types of local variables valid in the current subtree.
}

\FORALL{$i \in \{1, \dots, |startingNode.getChildren()|\}$ ordered from the lowest to the highest}
    \STATE $subnode = startingNode.getChild(i)$
    \STATE $analysisOfExpressionTypes(subnode, contextVarTypes)$ \COMMENT{recursion}
    \STATE $newVars = get(subnode, newVars)$    
    \STATE $contextVarTypes = mergeContextVarTypes(contextVarTypes, newVars)$
\ENDFOR
\IF{$is(startingNode, ExprNode)$}
    \STATE $memorizeType(startingNode, determineExpressionType(startingNode))$
\ENDIF
\IF{$is(startingNode, VariableBindingNode)$}
    \STATE $type = null$
    \STATE $typeNode = startingNode.getChild(TypeNode)$
    \IF{$typeNode \neq null$}
        \STATE $type = getTypeTN(typeNode)$
    \ELSE
        \STATE $type = determineExpressionType(startingNode.getChild(BindingSequenceNode).getChild(ExprNode), contextVarTypes)$
    \ENDIF
    \STATE $set(startingNode, newVars, (startingNode.varName, type))$
\ELSIF{$is(startingNode, TupleStreamNode)$}
    \FORALL{$varBindingNode \in startingNode.getChildren()$}
        \STATE $add(startingNode, newVars, get(varBindingNode, newVars))$
    \ENDFOR
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Function getVariableType}
\label{ALG_function_get_variable_type}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $getGlobalVariableType()$: Function returning a determined type of a specified global variable. \\
    $contextVarTypes$: Types of local variables valid in current context. \\
    $var$: A variable to determine the type of.
}
\ENSURE The type of variable $var$.

@@

The types of processed global variables are available trough function $getGlobalVariableType()$. If the function gets a variable that is not a global one, it returns $null$. Function $getVariableType()$ called in Algorithm \ref{ALG_func_determineExpressionType} is defined in Algorithm \ref{ALG_function_get_variable_type}. It checks if the specified variable is amongst the given local variables. If so, it returns its type, else it handles the variable as a global one and if such global variable does not exist the result is $null$.

\subsection{Determination of Expression Types}
To determine types of expressions, we use already introduced function $analysisOfExpressionTypes$. The starting node (its first argument) is the node representing the query body and the variable context is empty as there cannot be any local variable valid in the body node.

@@

If the operator is one of $<$, $>$, $<=$, $>=$, $=$, $!=$ (the class of the expression node is $Operator$ and it represents one of $GEN\_LESS\_THAN$, $GEN\_GREATER\_THAN$, $GEN\_LESS\_THAN\_EQUALS$, $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_EQUALS$, $GEN\_NOT\_EQUALS$), one operand is $E$ and the type $T$ of the other operand is one of built-in types, then $T$ is also the inferred type of $S$.

@@

\section{Step 4: Key Discovery}
In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in section \ref{Keys}, and extending it. Firstly, the syntax tree is searched for certain forms of FLWOR expressions to infer the keys from. Then, certain costructs are used to support or degrade the evidence of inferred statements in a final summarization.

@@

The two join patterns from the original approach are shown in Listings \ref{listing_for_join_pattern} and \ref{listing_let_join_pattern}. Additional join patterns are introduced in Listings \ref{listing_other_for_join_pattern}, \ref{listing_other_let_join_pattern} and \ref{listing_join_pattern_3}. In all join patterns $P_1$, $P_2$, $L_1$, $L_2$ are XQuery paths without predicates, using only child and descendant axes. Actually, the join patterns from \ref{listing_other_for_join_pattern} and \ref{listing_other_let_join_pattern} are covered by the for and let join pattern in the original approach, but we introduce them explicitly because their structure in the syntax tree is different.

@@

As described in the original method and summarized in section \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns, and rules to classify each occurrence into one of these cases. For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5. \todo{Popisat dovod nizsej vahy (nie je take jasne, ze je to par key-foreign key), pripadne zmenit vahu.}

@@

\STATE $bindingNodes = bindingNode \in flworNode.getChild(TupleStreamNode).getChildren()$

@@

\FORALL{$bindingNode \in bindingNodes$}
    \STATE $bindingExpr = bindingNode.getChild(BindingSequenceNode).getChild(ExprNode)$
    \IF{$getType(BindingExpr) == PathType$ using only child and descendant axes, ending with exactly one predicate $P$}
        \FORALL{$(var, node) \in forVars$}
            \STATE $forVars = determinateJoinPatterns(bindingNode, P, node, var, forVars)$
        \ENDFOR
    \ELSIF{$getType(BindingExpr) == PathType$ without predicates using only child and descendant axes}
        \IF{$is(bindingNode, ForClauseNode$}
            \STATE add $(bindingNode.varName, bindingNode)$ to forVars
        \ENDIF
    \ENDIF
\ENDFOR

@@

@@

\subsection{Rejection of Uniqueness}
While the previous steps of the key discovery produce positive statements about the keys, in this step, the algorithm searches for evidence on non-uniqueness of elements and attributes. An aim of this step is to eliminate or decrease the number of falsely inferred keys.
\todo[inline]{Napisat niekam, ze odvodzovanie kludov predpoklada, ze kazdy element/element join je robeny cez key-foreign key pair.}
When the algorithm discovers that a certain element is probably not unique, it assign the element with a negative weight that will decrease the sum of its weights in the summarizing step.

@@

Contrary the key discovery, $PathTypes$ utilized in this step may contain predicates. Predicates in a path restrict its result, and therefore, if a set of elements or attributes selected by a path with predicates is not unique, neither is a set of elements or attributes selected by the same path stripped of the predicates. \todo[inline]{Toto asi nebude uplne pravda pre nejake exoticke predikaty. Overit pri implementacii.}

@@

\IF{$is(node, FLWORExprNode)$}
    \FORALL{$bindingNode \in node.getChild(TupleStreamNode).getChildren()$}
        \IF{$is(bindingNode, ForClauseNode)$}
            \STATE $expr = bindingNode.getChild(BindingSequenceNode).getChild(ExprNode)$
            \IF{$getType(expr) == PathExpr$ without predicates, using only child and descendant axes}
                \STATE add $bindingNode.varName$ to $forVars$
            \ENDIF
        \ENDIF
    \ENDFOR
\ENDIF

>>>>

\begin{algorithm}
\caption{Processing of Functions}
\label{ALG_processing_of_functions}
\begin{algorithmic}[1]
\REQUIRE $syntaxTree$: A reference to the root node of a syntax tree.
\ENSURE An associative array of function names with their types and references to the their declaration nodes.

\STATE $functionArray$ = an empty associative array
\FORALL{$functionDeclarationNode \in$ \\ $getFunctionDeclarationNodes(syntaxTree)$}
    \STATE $functionName = functionDeclarationNode.funcName$
    \STATE $typeNode = functionDeclarationNode.getChild(TypeNode)$
    \STATE $functionArray[functionName] =$ \\ $(getTypeTN(typeNode), functionDeclarationNode)$
\ENDFOR
\RETURN $functionArray$
\end{algorithmic}
\end{algorithm}

@@

Later phases of the algorithm need to process the declarations of certain functions. Therefore, we also store references to the entire function declaration nodes, accessible using function $getFunctionDeclNode()$.

@@

Another group consists of functions $set()$, $add()$, and $get()$. Function $set()$ memorizes a given value of a specified property of a specified node. It is used to assign a node with a named value, for example, $set(someNode, color, 'red')$ will assign $someNode$ with string value $'red'$ wich can be then retrieved by function $get()$, specifying the node and the property name. For example,\linebreak $get(someNode, color)$ returns $'red'$. A subsequent call of $set()$ assigning a node with a value of already assigned property will overwrite it, leaving the property with the newer value. However, a property can have several values and this can be achieved using function $add()$ with the same syntax as function $set()$ but instead of overwriting the existing property, $add()$ will append the new value to the existing ones. Then, a call of $get()$ on this node and property returns a list of all its values, preserving the order of their addition.

@@

\begin{algorithm}
\caption{Function determineExpressionType()}
\label{ALG_func_determineExpressionType}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $exprNode$: An expression node in the syntax tree. \\
    $contextVarTypes$: Types of local variables valid in the current subtree.
}
\ENSURE The expression type.

@@

\begin{algorithm}
\caption{Function determineOperatorType()}
\label{ALG_func_determineOperatorType}
\begin{algorithmic}[1]
\REQUIRE{\ \\ 
    $operatorNode$: An operator expression node in the syntax tree. \\
    $contextVarTypes$: Types of local variables valid in the current subtree.
}
\ENSURE The operator expression type.

\STATE $type = null$
\STATE $operator = operatorNode.operator$
\IF{$isOperatorClassComparison(operator)$}
    \STATE $type = xs:boolean$, exactly-one cardinality
\ELSIF{$isOperatorClassAddition(operator)$}
    \STATE $leftOperandType = getType(exprNode.leftSide)$
    \STATE $rightOperandType = getType(exprNode.rightSide)$
    \IF{$isNumericType(leftOperandType)$ $AND$ $isNumericType(rightOperandType)$}
        \STATE $type = selectMoreGeneralNumericType(leftOperandType,$ $rightOperandType)$
    \ELSIF{$isNumericType(leftOperandType)$}
        \STATE $type = leftOperandType$
    \ELSIF{$isNumericType(rightOperandType)$}
        \STATE $type = rightOperandType$
    \ENDIF
\ENDIF

@@

The algorithm iterates through the variable declaration nodes from the prolog and if the variable is explicitly assigned with its type, the type is noted. Otherwise, an attempt of the type deduction of the binding expression is made. The deduction of the type from the expression is presented in Algorithm \ref{ALG_func_determineExpressionType}. On input, it takes an expression node and types of local variables that are valid in the expression (also called context of variables or variable context). The presented algorithm is just shortened illustration since the complete version is too long to be presented in this text. Nevertheless, the principle is straightforward. Depending on the class of the expression node (show in Figure \ref{FIG_syntax_tree_node_types_part_4}), its type can be either determined directly or it depends on its subexpressions. A part of the function's semantic is separated in Algorithm \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.

Function $determineExpressionType$ is called within function $analysis\-Of\-Expression\-Types$ presented in Algorithm \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions. Two important questions about this function emerge. How (in which order) are the tree nodes recursively processed and how does the function handle definitions (bindings) of new variables which may also appear in the expressions?

The order of the recursion is to firstly process children of a node and then the node itself. The order of child processing is specified by function $o$ from the definition of the syntax tree in Definition \ref{DEF_syntax_tree}. The reason is that an expression node may need to know types of its subexpressions to determine its own type and we want to determine types of all expressions. Therefore, using the described order, subexpressions of an expression are processed first, and then the expression itself without any need for further recursion, because the types of the subexpressions are already determined.

@@

\begin{algorithm}
\caption{Function analysisOfExpressionTypes}
\label{ALG_func_analysisOfExpressionTypes}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $startingNode$: A node determining a subtree to perform the analysis on. \\
    $contextVarTypes$: Types of local variables valid in the current subtree.
}

\FORALL{$i \in \{1, \dots, |startingNode.getChildren()|\}$ ordered from the lowest to the highest}
    \STATE $subnode = startingNode.getChild(i)$
    \STATE $analysisOfExpressionTypes(subnode, contextVarTypes)$ \COMMENT{recursion}
    \STATE $newVars = get(subnode, newVars)$    
    \STATE $contextVarTypes = mergeContextVarTypes(contextVarTypes, newVars)$
\ENDFOR
\IF{$is(startingNode, ExprNode)$}
    \STATE $memorizeType(startingNode, determineExpressionType(startingNode))$
\ENDIF
\IF{$is(startingNode, VariableBindingNode)$}
    \STATE $type = null$
    \STATE $typeNode = startingNode.getChild(TypeNode)$
    \IF{$typeNode \neq null$}
        \STATE $type = getTypeTN(typeNode)$
    \ELSE
        \STATE $type = determineExpressionType(startingNode.$ $getChild(BindingSequenceNode).getChild(ExprNode),$ $contextVarTypes)$
    \ENDIF
    \STATE $set(startingNode, newVars, (startingNode.varName, type))$
\ELSIF{$is(startingNode, TupleStreamNode)$}
    \FORALL{$varBindingNode \in startingNode.getChildren()$}
        \STATE $add(startingNode, newVars, get(varBindingNode, newVars))$
    \ENDFOR
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Function getVariableType}
\label{ALG_function_get_variable_type}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $contextVarTypes$: Types of local variables valid in current context. \\
    $var$: A variable to determine the type of.
}
\ENSURE The type of variable $var$.

@@

The types of processed global variables are available trough function \linebreak $getGlobalVariableType()$. If the function gets a variable that is not a global one, it returns $null$. Function $getVariableType()$ called in Algorithm \ref{ALG_func_determineExpressionType} is defined in Algorithm \ref{ALG_function_get_variable_type}. It checks if the specified variable is amongst the given local variables. If so, it returns its type, else it handles the variable as a global one and if such global variable does not exist the result is $null$.

\subsection{Determination of Expression Types}
To determine types of expressions, we use already introduced function \linebreak $analysisOfExpressionTypes$. The starting node (its first argument) is the node representing the query body and the variable context is empty as there cannot be any local variable valid in the body node.

@@

If the operator is one of $<$, $>$, $<=$, $>=$, $=$, $!=$ (the class of the expression node is $Operator$ and it represents one of $GEN\_LESS\_THAN$, \linebreak $GEN\_GREATER\_THAN$, $GEN\_LESS\_THAN\_EQUALS$,\linebreak $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_EQUALS$,\linebreak $GEN\_NOT\_EQUALS$), one operand is $E$ and the type $T$ of the other operand is one of built-in types, then $T$ is also the inferred type of $S$.

@@

\section{Step 4: Key Discovery}
In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in Chapter \ref{Keys}, and extending it. Firstly, the syntax tree is searched for forms of FLWOR expressions to infer the keys from. Then, as described later, certain costructs are used to support or degrade the evidence of inferred statements in a final summarization.

@@

The two join patterns from the original approach are shown in Listings \ref{listing_for_join_pattern} and \ref{listing_let_join_pattern}. Additional join patterns are introduced in Listings \ref{listing_other_for_join_pattern}, \ref{listing_other_let_join_pattern} and \ref{listing_join_pattern_3}. In all join patterns $P_1$, $P_2$, $L_1$, $L_2$ are XQuery paths without predicates, using only child and descendant axes. Actually, the join patterns from Listings \ref{listing_other_for_join_pattern} and \ref{listing_other_let_join_pattern} are covered by the for and let join pattern in the original approach, but we introduce them explicitly because their structure in the syntax tree is different.

@@

As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns, and rules to classify each occurrence into one of these cases. For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5. \todo{Popisat dovod nizsej vahy (nie je take jasne, ze je to par key-foreign key), pripadne zmenit vahu.}

@@

\STATE $bindingNodes = bindingNode \in flworNode.getChild(TupleStreamNode).$ $getChildren()$

@@

\FORALL{$bindingNode \in bindingNodes$}
    \STATE $bindingExpr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
    \IF{$getType(BindingExpr) == PathType$ using only child and descendant axes, ending with exactly one predicate $P$}
        \FORALL{$(var, node) \in forVars$}
            \STATE $forVars = determinateJoinPatterns(bindingNode, P, node, var,$ $forVars)$
        \ENDFOR
    \ELSIF{$getType(BindingExpr) == PathType$ without predicates using only child and descendant axes}
        \IF{$is(bindingNode, ForClauseNode$}
            \STATE add $(bindingNode.varName, bindingNode)$ to forVars
        \ENDIF
    \ENDIF
\ENDFOR

@@

\todo[inline]{Alg 6.12 FLWOR expressions - pre podmienky na riadkoch 13 a 17 definovat nejake funkcie}

@@

\subsection{Rejection of Uniqueness}
While the previous steps of the key discovery produce positive statements about the keys, in this step, the algorithm searches for evidence on non-uniqueness of elements and attributes. The aim of this step is to eliminate or decrease the number of falsely inferred keys.
\todo[inline]{Napisat niekam, ze odvodzovanie kludov predpoklada, ze kazdy element/element join je robeny cez key-foreign key pair.}
When the algorithm discovers that a certain element is probably not unique, it assign the element with a negative weight that will decrease the sum of its weights in the summarizing step.

@@

Contrary the key discovery, $PathTypes$ utilized in this step may contain predicates. Predicates in a path restrict its result, and therefore, if a set of elements or attributes selected by a path with predicates is not unique, neither is a set of elements or attributes selected by the same path stripped of the predicates. \todo[inline]{Toto asi nebude uplne pravda pre nejake exoticke predikaty. Overit pri implementacii. A: Kdyžtak popište.}

@@

\IF{$is(node, FLWORExprNode)$}
    \FORALL{$bindingNode \in node.getChild(TupleStreamNode).getChildren()$}
        \IF{$is(bindingNode, ForClauseNode)$}
            \STATE $expr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
            \IF{$getType(expr) == PathExpr$ without predicates, using only child and descendant axes}
                \STATE add $bindingNode.varName$ to $forVars$
            \ENDIF
        \ENDIF
    \ENDFOR
\ENDIF
